---
title: "Abdulaziz Alqumayzi G200007615 Module6_CT3"
author: "Rudra S Bandhu"
date: "11/13/2021"
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(dplyr)
library(caret)
library(ggplot2)
library(multcomp)
library(Metrics)
library(stats)
library(Metrics)
library(car)
library(jmv)
library(readxl)
```


## Import datasets 
```{r}

problem_1 <- read_excel("problems.xlsx", sheet = "problem 1")
problem_3 <- read_excel("problems.xlsx", sheet = "problem 3")
problem_4 <- read_excel("problems.xlsx", sheet = "problem 4")

```


## Problem 1

```{r}
# Code for problem 1

# Apply Anova test
model_p1 <- aov(test ~ temp, data = problem_1) 
summary(model_p1)
print(model_p1)

```

## Problem 1 Questions and Answers

*a* Specify an appropriate null hypothesis 

H0: The null hypothesis is that the mean of test at different temperatures is equal 

*b* Test the hypothesis that the polymer performs equally well at all three temperatures at 5 percent level of significance

At 5 percent level of significance. The p-value (0.289) is greater than 0.05 so we fail to reject the null hypothesis.

*c* Test the hypothesis that the polymer performs equally well at all three temperatures at 1 percent level of significance

At 1 percent level of significance. The p-value  (0.289) is greater than 0.1 so we fail to reject the null hypothesis.

*d* State your conclusion from the analysis.

We can conclude that the ability of the polymer to remove toxic wastes from water tests has no different at three different temperatures.


## Problem 2

```{r}
# Code for problem 2

gmean <- (32+40+30)/3 # grand mean 
SSTR<-(12*((32-gmean)^2+(40-gmean)^2+(30-gmean)^2)) # Sum of Squares treatment
MSTR<-SSTR/(3-1)
SSe<-(12-1)*(33+44+40) # Sum of Squares error
MSE<-SSe/(36-3)
F<-MSTR/MSE # f-test 
P <- 1-pf(8.6154,2,33) # p-value 
q<-3.48
MOE<-3.48*sqrt(MSE/12)

cat(paste("F-test = ",round(F, digits = 5), " p-value = ", round(P, digits = 5),'\n\n',
          'CI of A-B: (',round(32-40-MOE, digits = 5),round(32-40+MOE,digits = 5),')\n',
      'CI of A-C: (',round(32-30-MOE, digits = 5),round(32-30+MOE,digits = 5),')\n',
      'CI of B-C: (',round(40-30-MOE, digits = 5),round(40-30+MOE,digits = 5),')'))

```

## Problem 2 Questions and Answers

 *a* Test the hypothesis that the mean time to clear a mild asthmatic attack is the same for all three steroids. Use the 5 percent level of significance.
 
At 5 percent level of significance. The p-value (0.00098) is less than 0.05 so we have significant evidence to reject the null hypothesis and accept the alternative.
 
*b* Find confidence intervals for all differences of means ( ) that, with 95 percent confidence, are valid.

At 95% confidence: 

 Confidence interval of A-B: ( -14.27366 -1.72634 )
 Confidence interval of A-C: ( -4.27366 8.27366 )
 Confidence interval of B-C: ( 3.72634 16.27366 )


## Problem 3

```{r}
# Code for problem 3

#Apply Anova test
model_p3 <- aov(Yield ~ Area, data = problem_3)
summary(model_p3)
print(model_p3)

```
## Problem 3 Questions and Answers

*a* State your null hypothesis.

H0: The null hypothesis is that the mean of planting snow peas in different regions is equal. 

*b* What analysis strategy 

One-way independent ANOVA Test to compare several means when those means have come from different groups.

*c* Calculate the test statistic and p-value.

F-statistic value = 5.12 and P-value = 0.013

*d* Calculate the residual error.

Residual standard error = 1.257835

*e* State your conclusion.

The p-value (0.013) is less than 0.05 at a 5 percent level of significance. So we have sufficient evidence to reject the null. 

We can conclude that planting snow peas in different regions are different in different regions.

*f* What approach can you take to reduce the residual error. Discuss at least 2 ideas and
justify 

1- Reduce variability: The less that your data varies, the more precisely you can estimate a population parameter.That's because reducing the variability of your data decreases the standard deviation and, thus, the margin of error for the estimate. 

2- Elimination of confounds: Unmeasured variables confound the results. If any variables are known to impact the dependent variable being assessed, ANCOVA is an excellent way to eliminate their bias. Once a potential confounding variable is identified, it may be tested and included as a covariate in the analysis.

3- Increase the sample size: Often, the most practical way to decrease the margin of error is to increase the sample size. Usually, the more observations that you have, the narrower the interval around the sample statistic is. Thus, you can often collect more data to obtain a more precise estimate of a population parameter.

## Problem 4


```{r}
# Code for problem 4


problem_4$Area <- as.factor(problem_4$Area)  

model_p4_1 <- ancova(data = problem_4, dep =  Yield, factors = Area, covs = 'Height(inches)',postHoc = ~ Area, modelTest = T )
print(model_p4_1)

cat(paste("New Residual error = ",round(sqrt(40.168312/26), digits = 5),
      "\nOld Residual error = ", round(1.257835, digits = 5)))


```
 
*a* What analysis strategy would you use to take advantage of the additional data you have

The Analysis of Covariance (ANCOVA) is used to compare means of an outcome variable between two or more groups taking into accountvariability of other variables, called covariates.

*b*  What is the dependence of the yield on the height for each of the areas? Does it differ
significantly between the 3 areas. What can you conclude (max 10 sentences)

We can see the p-value of dependency between Yield on height is 0.2102391. Which indicates that there is a weak dependency.

Code Results:

   Area         Area    Mean Difference    SE           df          t             p-tukey     
 ──────────────────────────────────────────────────────────────────────────────────────────── 
   FS      -    PS           -0.5848511    0.6044905    26.00000    -0.9675109    0.6034443   
           -    SH           -0.1780560    1.0130785    26.00000    -0.1757574    0.9831254   
   PS      -    SH            0.4067951    1.2186513    26.00000     0.3338076    0.9405663   
   


We can see the all p-values of the test are less than 0.05 from the code results. Which indicate that there are nor deffirence between the dependences for each area.

*c*  Calculate the test statistic and p-value.

                     Sum of Squares    df    Mean Square    F            p           
 ─────────────────────────────────────────────────────────────────────────────────── 
   Overall model           4.092615     3      1.3642050    4.0455540    0.0173929
   
We can see from the Overall model results that: 

F-statistic value = 4.0455540 and P-value = 0.0173929

*d*  Calculate the residual error and compare that to what you found in problem 2.  

In general, the smaller the residual standard deviation/error, the better the model fits the data. 

New model residual error =  1.24295 
Old model residual error =  1.25784

We can see that the residuals are slightly near to each other. But in this new model, there is an indication that the new model does improve slightly which gives us more confidence in our decision in the previous model.



*e* Specify your final model.

We used ancova() function for analysis of covariance to test the model that contains one dependent variable, one factor variable, and one covariate variable.

The code used is the following: 

ancova(data = problem_4, dep =  Yield, factors = Area, covs = 'Height(inches)',postHoc = ~ Area, modelTest = T )

modelTest parameter used to show the overall model result.

The code result is the following: 

 ANCOVA

 ANCOVA - Yield                                                                      
 ─────────────────────────────────────────────────────────────────────────────────── 
                     Sum of Squares    df    Mean Square    F            p           
 ─────────────────────────────────────────────────────────────────────────────────── 
   Overall model           4.092615     3      1.3642050    4.0455540    0.0173929   
   Area                    1.542927     2      0.7714634    0.4993501    0.6126274   
   Height(inches)          2.549688     1      2.5496883    1.6503530    0.2102391   
   Residuals              40.168312    26      1.5449351                             
 ─────────────────────────────────────────────────────────────────────────────────── 


 POST HOC TESTS

 Post Hoc Comparisons - Area                                                                  
 ──────────────────────────────────────────────────────────────────────────────────────────── 
   Area         Area    Mean Difference    SE           df          t             p-tukey     
 ──────────────────────────────────────────────────────────────────────────────────────────── 
   FS      -    PS           -0.5848511    0.6044905    26.00000    -0.9675109    0.6034443   
           -    SH           -0.1780560    1.0130785    26.00000    -0.1757574    0.9831254   
   PS      -    SH            0.4067951    1.2186513    26.00000     0.3338076    0.9405663   
 ──────────────────────────────────────────────────────────────────────────────────────────── 
 
 
The residual error was calculated using the following code:

sqrt(40.168312/26)

*f* State your conclusion 

As we tested in the previous problem 3. The p-value of model 3 is 0.013 and the p-value of model 4 is 0.017. Both models indicate that we have sufficient evidence to reject the null and accept the alternative. Which was we can conclude that planting snow peas in different regions is different in different regions. Also, the decreased residual gives us more confidence in our previous decision. 


References: 

Field, A. P., Miles, J., &amp; Field Zoë. (2017). Discovering statistics using R. W. Ross MacDonald School Resource Services Library. 

Choueiry, G.(2021)Residual Standard Deviation/Error: Guide for Beginners. Retrieved November 13, 2021, from https://quantifyinghealth.com/residual-standard-deviation-error/

finnstats (2021) How to perform ANCOVA in R. Retrieved November 13, 2021, from https://www.r-bloggers.com/2021/07/how-to-perform-ancova-in-r/

